---
output: github_document
always_allow_html: yes
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.retina = 2
)
```

[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/0.1.0/active.svg)](http://www.repostatus.org/#active) 
[![Travis-CI Build Status](https://travis-ci.org/hrbrmstr/ggalt.svg?branch=master)](https://travis-ci.org/hrbrmstr/ggalt) 
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/hrbrmstr/ggalt?branch=master&svg=true)](https://ci.appveyor.com/project/hrbrmstr/ggalt) 
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/ggalt)](https://CRAN.R-project.org/package=ggalt)
[![downloads](http://cranlogs.r-pkg.org/badges/grand-total/ggalt)](https://CRAN.R-project.org/package=ggalt)

# `ggalt`

A compendium of 'geoms', 'coords', 'stats', scales and fonts for 'ggplot2', including splines, 1d and 2d densities, univariate average shifted histograms, a new map coordinate system based on the 'PROJ.4'-library and the 'StateFace' open source font 'ProPublica'.

## Installation

```{r eval=FALSE}
# you'll want to see the vignettes, trust me
install.packages("ggplot2")
install.packages("ggalt")
# OR: devtools::install_github("hrbrmstr/ggalt")
```

## Geoms/Stats

|  __Geom__ |  __Description__ | __Uses__ |
|:--|:--|:--|
| `geom_ubar()` | Uniform width bar charts |
| `geom_horizon()` | Horizon charts (modified from <https://github.com/AtherEnergy/ggTimeSeries>) |
| `coord_proj()` | Like `coord_map`, only better (prbly shld use this with `geom_cartogram` as `geom_map`'s new defaults are ugh)|
| `geom_xspline()` | Connect control points/observations with an X-spline|
| `stat_xspline()` | Connect control points/observations with an X-spline|
| `geom_bkde()` |	Display a smooth density estimate | `KernSmooth::bkde`
| `geom_stateface()`|	Use ProPublica's StateFace font in ggplot2 plots |
| `geom_bkde2d()` | Contours from a 2d density estimate. | `KernSmooth::bkde2D`
| `stat_bkde()` |	Display a smooth density estimate | `KernSmooth::bkde`
| `stat_bkde2d()` |	Contours from a 2d density estimate. | `KernSmooth::bkde2D`
| `stat_ash()` | Compute and display a univariate averaged shifted histogram (polynomial kernel) | `ash::ash1`/`ash::bin1`
| `geom_encircle()`|	Automatically enclose points in a polygon |
| `geom_lollipop()`| Dead easy lollipops (horizontal or vertical) |
| `geom_dumbbell()` | Dead easy dumbbell plots |
| `stat_stepribbon()` | Step ribbons |
| `annotation_ticks()` | Add minor ticks to identity, exp(1) and exp(10) axis scales independently of each other. |
| `geom_spikelines()` | Instead of geom_vline and geom_hline a pair of segments that originate from same c(x,y) are drawn to the respective axes. |
| `byte_format()`| helpers. e.g. turn `10000` into `10 Kb` |

__note__: plotly integration for a few of the ^^ geoms

### Code of Conduct

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). 
By participating in this project you agree to abide by its terms.
