---
output:
  html_document:
    keep_md: true
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
fig.retina = 2,
  fig.path = "README_figs/README-"
)
```
[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/0.1.0/active.svg)](http://www.repostatus.org/#active) 
[![Travis-CI Build Status](https://travis-ci.org/hrbrmstr/ggalt.svg?branch=master)](https://travis-ci.org/hrbrmstr/ggalt) 
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/ggalt)](http://cran.r-project.org/web/packages/ggalt) 
![downloads](http://cranlogs.r-pkg.org/badges/grand-total/ggalt)

`ggalt` : Extra Coordinate Systems, Geoms and Statistical Transformations for 'ggplot2'

A compendium of 'geoms', 'coords' and 'stats' for 'ggplot2', including splines, 1d and 2d densities, univariate average shifted histograms and a new map coordinate system based on the 'PROJ.4'-library.

The first three forays into this brave, new `ggplot2` world are _splines_! and being able to use the (much better) `KernSmooth::bkde` and `KernSmooth::bkde2D` for density plots and an initial port of the (still needing work) `coord_proj`.

The following functions are implemented:

- `coord_proj` : Like `coord_map` only better 😜
- `geom_xspline` : Connect control points/observations with an X-spline
- `stat_xspline` : Connect control points/observations with an X-spline
- `geom_bkde` :	Display a smooth density estimate (uses `KernSmooth::bkde`)
- `stat_bkde` :	Display a smooth density estimate (uses `KernSmooth::bkde`)
- `geom_bkde2d` :	Contours from a 2d density estimate. (uses `KernSmooth::bkde2D`)
- `stat_bkde2d` :	Contours from a 2d density estimate. (uses `KernSmooth::bkde2D`)
- `stat_ash` : Compute and display a univariate averaged shifted histogram (polynomial kernel) (uses `ash::ash1`/`ash::bin1`)

### News

- Version 0.1.1 released - CRAN!
- Version 0.1.0.9000 released - Tweaks for ggplot2 2.0 release
- Version 0.0.4.9000 released - `stat_ash`
- Version 0.0.3.9000 released - `coord_proj`! (requires my github copy of ggplot2 for now)
- Version 0.0.2.9005 released - cleanup before blog post
- Version 0.0.2.9002 released - working 2D density plots
- Version 0.0.2.9000 released

### Installation

```{r eval=FALSE}
# you'll want to see the vignettes, trust me
install.packages("ggplot2")
install.packages("ggalt")
# OR: devtools::install_github("hrbrmstr/ggalt")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
options(width=120)
```

### Usage

```{r}
library(ggplot2)
library(gridExtra)
library(ggalt)

# current verison
packageVersion("ggalt")

set.seed(1492)
dat <- data.frame(x=c(1:10, 1:10, 1:10),
                  y=c(sample(15:30, 10), 2*sample(15:30, 10), 3*sample(15:30, 10)),
                  group=factor(c(rep(1, 10), rep(2, 10), rep(3, 10)))
)
```

### Splines!

```{r splines}
ggplot(dat, aes(x, y, group=group, color=group)) +
  geom_point() +
  geom_line()

ggplot(dat, aes(x, y, group=group, color=factor(group))) +
  geom_point() +
  geom_line() +
  geom_smooth(se=FALSE, linetype="dashed", size=0.5)

ggplot(dat, aes(x, y, group=group, color=factor(group))) +
  geom_point(color="black") +
  geom_smooth(se=FALSE, linetype="dashed", size=0.5) +
  geom_xspline(size=0.5)

ggplot(dat, aes(x, y, group=group, color=factor(group))) +
  geom_point(color="black") +
  geom_smooth(se=FALSE, linetype="dashed", size=0.5) +
  geom_xspline(spline_shape=-0.4, size=0.5)

ggplot(dat, aes(x, y, group=group, color=factor(group))) +
  geom_point(color="black") +
  geom_smooth(se=FALSE, linetype="dashed", size=0.5) +
  geom_xspline(spline_shape=0.4, size=0.5)

ggplot(dat, aes(x, y, group=group, color=factor(group))) +
  geom_point(color="black") +
  geom_smooth(se=FALSE, linetype="dashed", size=0.5) +
  geom_xspline(spline_shape=1, size=0.5)

ggplot(dat, aes(x, y, group=group, color=factor(group))) +
  geom_point(color="black") +
  geom_smooth(se=FALSE, linetype="dashed", size=0.5) +
  geom_xspline(spline_shape=0, size=0.5)

ggplot(dat, aes(x, y, group=group, color=factor(group))) +
  geom_point(color="black") +
  geom_smooth(se=FALSE, linetype="dashed", size=0.5) +
  geom_xspline(spline_shape=-1, size=0.5)
```

#### Alternate (better) density plots

```{r bkde_ash}
# bkde

data(geyser, package="MASS")

ggplot(geyser, aes(x=duration)) + 
  stat_bkde(alpha=1/2)

ggplot(geyser, aes(x=duration)) +
  geom_bkde(alpha=1/2)

ggplot(geyser, aes(x=duration)) + 
  stat_bkde(bandwidth=0.25)

ggplot(geyser, aes(x=duration)) +
  geom_bkde(bandwidth=0.25)

set.seed(1492)
dat <- data.frame(cond = factor(rep(c("A","B"), each=200)), 
                   rating = c(rnorm(200),rnorm(200, mean=.8)))

ggplot(dat, aes(x=rating, color=cond)) + geom_bkde(fill="#00000000")

ggplot(dat, aes(x=rating, fill=cond)) + geom_bkde(alpha=0.3)

# ash

set.seed(1492)
dat <- data.frame(x=rnorm(100))
grid.arrange(ggplot(dat, aes(x)) + stat_ash(),
             ggplot(dat, aes(x)) + stat_bkde(),
             ggplot(dat, aes(x)) + stat_density(),
             nrow=3)

cols <- RColorBrewer::brewer.pal(3, "Dark2")
ggplot(dat, aes(x)) + 
  stat_ash(alpha=1/3, fill=cols[3]) + 
  stat_bkde(alpha=1/3, fill=cols[2]) + 
  stat_density(alpha=1/3, fill=cols[1]) + 
  geom_rug() +
  labs(x=NULL, y="density/estimate") +
  scale_x_continuous(expand=c(0,0)) +
  theme_bw() +
  theme(panel.grid=element_blank()) +
  theme(panel.border=element_blank())
```

### Alternate 2D density plots

```{r bkde2d}
m <- ggplot(faithful, aes(x = eruptions, y = waiting)) +
       geom_point() +
       xlim(0.5, 6) +
       ylim(40, 110)

m + geom_bkde2d(bandwidth=c(0.5, 4))

m + stat_bkde2d(bandwidth=c(0.5, 4), aes(fill = ..level..), geom = "polygon")

```

### `coord_proj` LIVES! (still needs work)

```{r coord_proj}
world <- map_data("world")
world <- world[world$region != "Antarctica",]

gg <- ggplot()
gg <- gg + geom_map(data=world, map=world,
                    aes(x=long, y=lat, map_id=region))
gg <- gg + coord_proj("+proj=wintri")
gg
```

### Code of Conduct

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). 
By participating in this project you agree to abide by its terms.
